---
output: html_document
---


## Reading the input files and creating a clean data set to work with
The input data frame contains ***114 observations*** and ***38 variables***.

We wrote a function called **tidydata** that takes in a filename (with the path of the input file) and returns a clean data frame that we can start analyzing.

This function is stored as a script file named "tidydata.R" in our working directory that we need to source and run before we start the analysis.

```{r}
#Function that takes the survey file and returns a clean & tidy dataframe
tidydata<-function(filename)
{
  library(xlsx) 
  df<-read.xlsx(filename, sheetIndex = 1, header = T, stringsAsFactors = F)
  removeCol<-c(19:38,5:11)
  #remove unused columns
  df<-df[,-removeCol]
  #regroup columns logic-wise
  df<-df[c(1,2,5,6,4,7:11,3)]
  
  #clean up waiting list info
  df[,1]<-replace(df[,1],df[,1]=="Yes",1)
  df[,1]<-replace(df[,1],df[,1]=="No",0)
  df[,1] <- factor(df[,1], levels = c(0,1), labels = c("No","Yes"))
  names(df)[1] <- "waiting_list"
  
  #clean up Program column
  df[,2]<-replace(df[,2],df[,2]=="Ms in ds", "IDSE (master)")
  df[,2]<-replace(df[,2],df[,2]=="MSDS", "IDSE (master)")
  df[,2]<-replace(df[,2],df[,2]=="Data Science", "DS Certification")
  df[,2]<-replace(df[,2],df[,2]=="Data Science Certification", "DS Certification")
  df[,2]<-replace(df[,2],df[,2]=="QMSS", "Other masters")
  df[,2]<-replace(df[,2],df[,2]=="QMSS (master)", "Other masters")
  df[,2]<-replace(df[,2],df[,2]=="Applied Math", "Other masters")
  df[,2]<-replace(df[,2],df[,2]=="Ph.D.", "Other PhD")
  df[,2]<-replace(df[,2],df[,2]=="PhD Biomedical Informatics", "Other PhD")
  df[,2]<-factor(df[,2])
  names(df)[2] <- "program"
  
  #clean up gender
  df[,3]<-replace(df[,3],df[,3]=="she/her","F")
  df[,3]<-replace(df[,3],df[,3]=="he/him","M")
  df[,3]<-replace(df[,3],df[,3]=="doesn't matter","NA")
  df[,3]<-replace(df[,3],is.na(df[,3]),"NA")
  df[,3]<-factor(df[,3])
  names(df)[3] <- "gender"
  
  #clean up Text Editors
  df[,4]<-factor(df[,4])
  levels(df[,4])<-c(levels(df[,4]),"Eclipse","TextWrangler","None","Any","Jupyter")
  df[grep("Sublime Text / Eclipse",df[,4], ignore.case = FALSE),4]<- "Eclipse"
  df[grep("Sublime",df[,4], ignore.case = TRUE),4]<-"Sublime"
  df[grep("wrangler",df[,4], ignore.case = TRUE),4]<-"TextWrangler"
  df[grep("eclipse",df[,4], ignore.case = TRUE),4]<-"Eclipse"
  df[grep("haven't used any",df[,4], ignore.case = TRUE),4]<-"None"
  df[grep("20 years ",df[,4], ignore.case = TRUE),4]<-"Any"
  df[grep("Jupyter",df[,4], ignore.case = TRUE),4]<-"Jupyter"
  df[grep("ipynb",df[,4], ignore.case = TRUE),4]<-"Ipython"
  df[,4]<-droplevels(df[,4])
  names(df)[4] <- "text_editor"
  
  #Clean up Experience with tools
  
  #Split the column into a list of lists
  df$Experiences.with.tools = factor(df$Experiences.with.tools)
  temp<-strsplit(as.character(df$Experiences.with.tools),", ")
  #find the number of unique elements in the Experience with tools column
  skills<-unique(unlist(temp))
  #create a temporary data frame with 1 column for each unique skill
  newColNames<-skills
  skillsdf<-as.data.frame(matrix(nrow=nrow(df), 
                                 ncol=length(newColNames),
                                 dimnames=list(NULL, newColNames)))
  
  #The below for loop goes through each row in the excel file and looks at the
  #experience with tools column. If a skill is present, then you put a 1 against that
  #row and column in the temporary skills dataframe.
  for (i in 1:nrow(df))
  {
    skillList<-temp[[i]]
    for (j in 1:length(skillList))
    {
      skill<-skillList[j]
      skillsdf[i,skill]<-1
    }
  }
  #bind the temporary dataframe to the original dataframe
  cleandf<-cbind(df,skillsdf)
  return(cleandf)
}
```

The **tidydata** function does the following:

1. Removes the columns that have no observations in them.

2. Regrouped the columns in a logical manner.

3. Organized the factor levels in the following columns
  + Waiting List Info
  + Gender
  + Program
  + Text Editors

4. We split the "Experience with Tools"" column to multiple columns with a 1/0 value depending on if the skill was reported or not. Splitting the variable makes it tidy.


```{r message = FALSE}
#assuming that you have set your working directory to where you downloaded the input data
filename<-"Survey+Response.xlsx"
source("tidydata.R")
df<-tidydata(filename)
```
Now this data frame **df** has the data we want to start analyzing.


## Exploratory Plot #1 
###Distribution of class by Gender & Program
We broke down the participants of the class by gender and their program affiliation. The bar plot shows that the majority of our class either belongs to IDSE (master) or DS Certification. We have a significant representation from Statistics (mster). We also note that majority of students are male.

```{r}
GenderProgramPlot<-function(d)
{
  library(ggplot2)
  #Plotting stacked bar plot to show distribution of students in class by program and gender
  
  #Create a table with counts
  pdata<-data.frame(table(d$gender, d$program))
  names(pdata)<-c("Gender","Program", "Count")
  
  #format the x tick labels so that they wrap to the next line and not bleed
  levels(pdata$Program)<-gsub(" ","\n", levels(pdata$Program))
  
  #ggplot code
  ggp<-ggplot(data=pdata, aes(x=reorder(Program,-Count), y=Count, fill = Gender)) + 
      geom_bar(stat = "identity") +
      theme_bw() +
      ggtitle("Distribution of class by Gender and Program") +
      xlab("Program") +
      ylab("Number of students in Program")
  return(ggp)
}

GenderProgramPlot(df)
```

## Exploratory Plot #2
###Text Editor use by program (histogram)
We were curious about what text editors were most commonly used by program affiliation. We took our 4 most representative programs and plotted a histogram of counts broken down by programs.

```{r}
TextEditorProgramPlot<-function(d)
{
  library(ggplot2)
  #Plotting stacked bar plot to show distribution of students in class by program and gender
  
  #Create a table with counts
  te<-data.frame(table(d$text_editor, d$program))
  names(te)<-c("TextEditor","Program", "Count")
  te<-te[ which(te$Program!='Other PhD'),]
  
  #format the LEGEND labels they wrap to the next line and not bleed
  levels(te$Program)<-gsub(" ","\n", levels(te$Program))
  levels(te$TextEditor)<-gsub(" ","\n", levels(te$TextEditor))
  levels(te$TextEditor)<-gsub("Webstorm,\npycharm","Webstorm/pycharm",levels(te$TextEditor))
  #ggplot code
  ggp<-ggplot(data=te, aes(x=reorder(TextEditor,-Count), y=Count, fill = Program)) + 
    facet_grid(~ Program)+
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(text = element_text(size = 7),axis.text.x = element_text(angle=90, vjust=1)) +
    ggtitle("Distribution of Text Editors Used by Program") + 
    xlab("Text Editor") +
    ylab("Counts")
  return(ggp)
}
TextEditorProgramPlot(df)
```

